---
title: "RevisionsOfR"
author: "Matheus Erthal"
date: "07/03/2015"
output: html_document
---

Revisions of R
==============

Here a try to keep records of nice functions and techniques learned in R.

About the cars dataset:

```{r, echo = FALSE}
summary(cars)
plot(cars)
```

Aggregate function creates a new dataset splitting the data into subsets and applying a funtion for each (e.g. mean).

```{r, echo=TRUE}
cars_2 <- aggregate(speed ~ dist, data = cars, FUN = mean)
plot(dist ~ speed, data=cars_2)
```

Number of levels from a column
```{r}
nlevels(as.factor(data$date)
```

Unique function
```{r}
unique(data$date)
```



##Formatting Data

Working with dates.

To get the current date:
```{r, echo=TRUE}
d2 <- Sys.Date()
class(d2)
```

###Function *format*

Formatting dates parameters:
%d = day as number (0-31)
%a = abbreviated weekday
%A = unabbreviated weekday
%m = month (00-12)
%M = unabbreviated month
%y = 2 digit year
%Y = 4 digit year

```{r, echo=TRUE}
format(d2, "%a %b %d")
```

Creating dates:

```{r}
x <- c("1jan1960", "2jan1960", "30jul1960")
z <- as.Date(x, "%d%b%Y")
z
```
  
Time difference of -1 days:
```{r}
z[1] - z[2]
as.numeric(z[1] - z[2])
```
  
Converting to Julian:
```{r}
weekdays(d2)
months(d2)
julian(d2)
```

A date can be converted to numeric and be converted again to date, if you use the origin:
```{r, echo=TRUE}
z <- as.numeric(z)
z <- as.Date(z, origin)
```

When having a column of the dataset with the dates and another with the intervals in seconds, an approach that can be adopted is:
```{r}
data$datetime <- as.POSIXct(data$date)
data$datetime <- data$datetime - 3600*21 + data$interval*60
```
In the example above, when converted the date to POSIXct all the times were 21h, so we have subtracted it before adding the interval in min.

The difference between POSIXct and POSIXlt is that the former stores seconds since UNIX epoch (+some other data) and the latter stores a list of day, month, year, hour, minute, second, etc.

strptime is a function to directly convert character vectors (of a variety of formats) to POSIXlt format.



###Package Lubridate

A simpler way to deal with data. The functions signatures stands for the formatation of the data passed as parameter, e.g. "mdy" stands for "month-day-year", "ymd_hms" stands for "year-month-day hour:minutes:seconds".

```{r, echo=TRUE}
library(lubridate)
ymd("20140108")
mdy("08/04/2013")
dmy("03-04-2013")
ymd_hms("2011-08-03 10:15:03", tz = "Pacific/Auckland")
```

Information about timezones: 
  
```{r, message=TRUE}
?Sys.timezone
```

Some functions have a sintax slightly different. Such as:

```{r, echo=TRUE}
x <- dmy(c("1jan2013", "2jan2013", "30jul2013"))
wday(x[1])
wday(x[1], label = TRUE)
```

Ultimately I want date and time as class *Date* or *POSIXct*, *POSIXlt*.
